---
title: "bulk-rnaseq-gallagher"
output: html_document
date: "2025-01-07"
---
#I don't know what this is. is it needed??
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Note:
I am doing the analysis on the R server from class. All of the files are uploaded there and it seems like it would have taken a long time to download all of the files I would have needed. I'll try again when I'm done there. This will just be a copy and paste of my code but I won't be able to run it from my local computer because the files are not there.

Before code gets put in here I will have made sure that it works on the R server though.


#QC 
```{r}
#run fastqc on the raw fastq files
fastqcr::fastqc(fq.dir = "~/data/bulk_RNA_seq/fastq_files/raw",
                qc.dir = "/home/willettb/bulk-rnaseq-gallagher/fastq_raw_qc",
                threads = 1,
                fastqc.path = "~/software/FastQC/fastqc")

#multiqc to check qc data
system(command = 'multiqc /home/willettb/bulk-rnaseq-gallagher/fastq_raw_qc --filename *fastqc.html --outdir /home/willettb/bulk-rnaseq-gallagher')

```

